---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# econtools

<!-- badges: start -->
![R CMD Check](https://github.com/tidy-intelligence/econtools/actions/workflows/R-CMD-check.yaml/badge.svg)
![Lint](https://github.com/tidy-intelligence/econtools/actions/workflows/lint.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/tidy-intelligence/econtools/graph/badge.svg)](https://app.codecov.io/gh/tidy-intelligence/econtools)
<!-- badges: end -->

The goal of econtools is to ...

## Installation

You can install the development version of econtools from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tidy-intelligence/econtools")
```

## Usage

This is a basic example which shows you how to solve a common problem:

```{r example}
library(econtools)
```

```{r}
df <- data.frame(
  geography_id = rep("USA", 5),
  year = 2019:2023,
  gdp = c(2.15e13, 2.14e13, 2.37e13, 2.60e13, 2.77e13)
)
```

Add most recent population number:
```{r}
add_population_column(df, id_column = "geography_id")
```

Add population by year:

```{r}
add_population_column(df, id_column = "geography_id", date_column = "year")
```

Similarly, for poverty ratio:
```{r}
add_poverty_ratio_column(df, id_column = "geography_id", date_column = "year")
```

Create a new column that calculates a value relative to the population, for instance gdp per capita:
```{r}
add_population_share_column(df, id_column = "geography_id", date_column = "year", value_column = "gdp")
```

Add income levels via:
```{r}
add_income_level_column(df, id_column = "geography_id")
```