---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# econtools

<!-- badges: start -->
![R CMD Check](https://github.com/tidy-intelligence/econtools/actions/workflows/R-CMD-check.yaml/badge.svg)
![Lint](https://github.com/tidy-intelligence/econtools/actions/workflows/lint.yaml/badge.svg)
[![Codecov test coverage](https://codecov.io/gh/tidy-intelligence/econtools/graph/badge.svg)](https://app.codecov.io/gh/tidy-intelligence/econtools)
<!-- badges: end -->

The goal of econtools is to provide tools for analyzing economic data similar to the Python [bblocks](https://github.com/ONEcampaign/bblocks/tree/main) package.

The package is part of the [EconDataverse](https://www.econdataverse.org/) family of packages aimed at helping economists and financial professionals work with sovereign-level economic data.

## Installation

You can install the development version of econtools from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("tidy-intelligence/econtools")
```

## Usage

This is a basic example which shows you how to solve a common problem:

```{r example}
library(econtools)
```

```{r}
df <- data.frame(
  id = rep("USA", 5),
  name = c("United States", "United.states", "US", "USA", "United States"),
  year = 2019:2023,
  gross_domestic_product = c(2.15e13, 2.14e13, 2.37e13, 2.60e13, 2.77e13)
)

add_population_column(df, id_column = "name", id_type = "regex")
```

The simplest way to add additional information is using an ISO 3166-1 alpha-3 code.

Add most recent population number:
```{r}
add_population_column(df, id_column = "id")
```

Add population by year:

```{r}
add_population_column(df, id_column = "id", date_column = "year")
```

Similarly, for poverty ratio:
```{r}
add_poverty_ratio_column(df, id_column = "id", date_column = "year")
```

Create a new column that calculates a value relative to the population, for instance GDP per capita:
```{r}
add_population_share_column(
  df,
  id_column = "id",
  date_column = "year",
  value_column = "gross_domestic_product"
)
```

Add income levels via:

```{r}
add_income_level_column(df, id_column = "id")
```

If you want to use another column and automatically map identifiers to a new `entity_id` column using the `econid` package:

```{r}
add_population_column(df, id_column = "name", id_type = "regex")
```

If you only want to add ISO-3 codes:

```{r}
add_iso3_codes_column(df, "name")
```

You can also add a column with standardized names:

```{r}
add_short_names_column(df, "name")
```

Finally, you can add columns from IMF WEO data such as the GDP in national currency

```{r}
add_gdp_column(df, id_column = "id", date_column = "year", usd = FALSE)
```

Or the government expenditure (only availbale in national currency): 

```{r}
add_gov_exp_column(df, id_column = "id", date_column = "year")
```

And share of government expenditure to GDP: 

```{r}
add_gov_exp_share_column(df, id_column = "id", date_column = "year")
```
